---
title: "Assigment 8 The Scriptors"
author: "Jelle ten Harkel & Thijs van Loon"
date: "19 januari 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### The needed libraries
```{r include=TRUE, message=FALSE}
library(raster)
library(RColorBrewer)
library(knitr)
source("R/Functions.R")
```


```{r load data, include=FALSE}
load("data/GewataB1.rda")
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")
load("data/trainingPoly.rda")
```

```{r calculations, include=FALSE}
gewata <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
gewata <- calc(gewata, fun=function(x) x / 10000)

vcfGewata[vcfGewata > 100] <- NA
vcfGewata[vcfGewata < 0] <- NA

covs <- addLayer(gewata, vcfGewata)
names(covs) <- c("band1","band2","band3","band4","band5","band7","VCF")
df <- as.data.frame(getValues(covs))
```

#### Analysis of the scatterplots  
```{r scatterplots, echo=FALSE,warning=FALSE}
par(mfrow=c(2,3))

plot(GewataB1,vcfGewata, col="blue")
plot(GewataB2,vcfGewata, col="green")
plot(GewataB3,vcfGewata, col="red")
plot(GewataB4,vcfGewata, col="purple")
plot(GewataB5,vcfGewata, col="orange")
plot(GewataB7,vcfGewata, col="grey")
```

#### Summary of our model
As can be seen band 7 has a low significance, it has been removed from the model

```{r linear model, echo=FALSE}
#Create the model
LMmodel <- lm(VCF ~ band1 + band2 + band3 + band4 + band5 + band7, data = df)
summary(LMmodel)
## We eliminate band 7, because low significance
LMmodel <- lm(VCF ~ band1 + band2 + band3 + band4 + band5, data = df)
```  


#### Plot our predicted VCF and the original VCF
```{r calculate, echo=FALSE}
## Predict tree cover
par(mfrow = c(1,2), oma=c(0,0,0,1))
predTC <- predict(covs,model=LMmodel, na.rm=TRUE)
predTC[predTC > 100] <- NA
predTC[predTC < 0] <- NA
plot(predTC, main="The predicted VCF",legend=FALSE)
plot(vcfGewata, main="The original VCF")
```

#### Summary  
```{r, echo=FALSE, warning=FALSE}
plot(vcfGewata,predTC,col="darkgreen",xlab="Original VCF [%]", ylab="Predicted VCF [%]")
```

Looking at the scatterplot between the original and predicted VCF, we can see that the prediction overall is quite correct. 
The cloud looks to be on a angle of 45 degrees. Therefore there is a relatively high correlation. 
Low values of the original VCF are harder to predict, and higher values are easier to get right. 
There are only a few locations with a VCF of lower than 5%  


```{r make dataframes, include=FALSE}
predTC_df <- as.data.frame(predTC)
vcfGewata_df <- as.data.frame(vcfGewata)
```
```{r calculate rmse, echo=FALSE}
rmse <- rmse2(vcfGewata_df,predTC_df)
```

```{r, include=FALSE}
## calculate rmse per landtype
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
classes <- rasterize(trainingPoly,covs, field='Code')

rmse_PP <- rmse2(vcfGewata,predTC)
rmse_PC <- zonal(rmse_PP,classes,fun="mean")

## drop unnecessary column
rmse_PC <- rmse_PC[,2]
## make it a matrix again
rmse_PC <- matrix(unlist(rmse_PC), ncol = 1, byrow = TRUE)
rmse_PC <- rbind(rmse_PC,rmse)
## change column names
rownames(rmse_PC) <-c("cropland", "forest", "wetland","overall")
colnames(rmse_PC) <- ("RMSE")
```
  
#### All the RMSE values  
```{r, echo=FALSE,results='asis'}
kable(rmse_PC)
```

In this table, we see that the overall RMSE is higher than the RMSE of cropland and forest. Only the RMSE of 
wetland higher. We also see that the RMSE of forest is roughly half the size of the overall RMSE, which is good thing
because VCF is a indicator for forest.